\tit

### 2021-02-28 paper.tex
tune margins for footnotes
a little more on proof relevance

\ifacm
\documentclass[acmsmall,screen,anonymous,timestamp,review]{acmart}
\else
\documentclass[hidelinks]{article}  % fleqn,12pt
\usepackage[margin=0.9in]{geometry}  % 0.12in, 0.9in, 1in
\usepackage[useregional]{datetime2}
\fi

@@

This paper makes the following specific contributions:
\begin{itemize}
\item
  A collection of lemmas are stated and proved about functions over lists, capturing the semantic essence of Brzozowski's syntactic technique of derivatives of regular expressions (\secref{Decomposing Languages}).
\item
  A duality is revealed between regular expressions and tries (prefix trees) in the context of language differentiation (\secref{From Languages to Parsing}).
  Each implementation contains exactly the same code but in transposed forms.
  Both are corollaries of the semantic lemmas and are automatically proved correct by type-checking.
\item
  The theory and algorithms are specified in terms of proof isomorphism rather than mere logical equivalence.
  As such, the differentiation lemmas for language concatenation and Kleene closure are more informative than in previous work, capturing \emph{all} parsings, including closure of languages containing the empty string.
\item
  The vocabulary of languages is generalized to predicates over arbitrary types, and then re-specialized to the usual vocabulary (\secref{Predicate Algebra}).
  This generalization consists of two useful covariant applicative functors on functions (the usual one on functions-from-\AF{X} and another on functions-to-\APT{Set}).
  % Properties are proved in terms of this more general vocabulary.
\item
  Properties are stated first at the level of predicates, making them easy to understand and prove, thanks to elimination of operational details.
  These properties are then transported effortlessly to the computable parsing implementations (\secref{Transporting Properties}).
\end{itemize}
Although many proofs are elided below, the source code for this paper is freely available and contains fully verified Agda code with no postulates.
All type signatures and definitions in the paper are generated by the Agda compiler.

>>>>

\ifacm
\documentclass[acmsmall,screen,anonymous,timestamp,review]{acmart}
%% \documentclass[acmsmall,screen,timestamp,nonacm]{acmart}
%% \authorsaddresses{}
\else
\documentclass[hidelinks]{article}  % fleqn,12pt
\usepackage[margin=1.3in]{geometry}  % 0.12in, 0.9in, 1in
\usepackage[useregional]{datetime2}
\fi

@@

This paper makes the following specific contributions:
\begin{itemize}
\item
  A collection of lemmas are stated and proved about functions over lists, capturing the semantic essence of Brzozowski's syntactic technique of derivatives of regular expressions (\secref{Decomposing Languages}).
\item
  A duality is revealed between regular expressions and tries (prefix trees) in the context of language differentiation (\secref{From Languages to Parsing}).
  Each implementation contains exactly the same code but in transposed forms.
  Both are corollaries of the semantic lemmas and are automatically proved correct by type-checking.
\item
  The theory and algorithms are specified in terms of proof isomorphism rather than the usual, coarser relation of logical equivalence.
  As such, the lemmas for language concatenation and Kleene closure are more informative than in previous work, capturing \emph{all} parsings distinctly, including closure of languages containing the empty string.
\item
  The vocabulary of languages is generalized from predicates on lists to predicates on arbitrary types, and then re-specialized to the usual vocabulary (\secref{Predicate Algebra}).
  This generalization consists of two useful covariant applicative functors on functions (the usual one on functions-from-\AF{X} and another on functions-to-\APT{Set}).
  % Properties are proved in terms of this more general vocabulary.
\item
  Properties are stated first at the level of predicates, making them easy to understand and prove, thanks to elimination of operational details.
  These properties are then transported effortlessly to the computable parsing implementations (\secref{Transporting Properties}).
\end{itemize}
Although many proofs are elided below, the source code for this paper is freely available and contains fully verified Agda code with no postulates.
All type signatures and definitions in the paper are generated by the Agda compiler.
